My title

### 2024-11-26 paper/paper.qmd
draft

[Overview paragraph

@@

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....]

@@

The dataset contains key variables spanning musical features, environmental data, and outcomes. Below is a summary of numerical variables:
```{r}
#| label: tbl-summary
#| tbl-cap: Summary Statistics
#| echo: false
#| warning: false
#| fig-align: center
summary_table <- analysis_data %>%
  reframe(
    Variable = c("Valence", "Mean Temp", "Danceability", "Energy", "Acousticness", "Tempo"),
    Mean = c(
      mean(Valence, na.rm = TRUE),
      mean(`Mean Temp`, na.rm = TRUE),
      mean(Danceability, na.rm = TRUE),
      mean(Energy, na.rm = TRUE),
      mean(Acousticness, na.rm = TRUE),
      mean(Tempo, na.rm = TRUE)
    ),
    SD = c(
      sd(Valence, na.rm = TRUE),
      sd(`Mean Temp`, na.rm = TRUE),
      sd(Danceability, na.rm = TRUE),
      sd(Energy, na.rm = TRUE),
      sd(Acousticness, na.rm = TRUE),
      sd(Tempo, na.rm = TRUE)
    ),
    Min = c(
      min(Valence, na.rm = TRUE),
      min(`Mean Temp`, na.rm = TRUE),
      min(Danceability, na.rm = TRUE),
      min(Energy, na.rm = TRUE),
      min(Acousticness, na.rm = TRUE),
      min(Tempo, na.rm = TRUE)
    ),
    Max = c(
      max(Valence, na.rm = TRUE),
      max(`Mean Temp`, na.rm = TRUE),
      max(Danceability, na.rm = TRUE),
      max(Energy, na.rm = TRUE),
      max(Acousticness, na.rm = TRUE),
      max(Tempo, na.rm = TRUE)
    )
  ) %>%
  mutate(
    Mean = round(Mean, 3),
    SD = round(SD, 3),
    Min = round(Min, 3),
    Max = round(Max, 3)
  )
kable(summary_table, format = "pipe")
```
These metrics reveal diverse song features and moderate variability in environmental temperature, offering rich explanatory potential for Valence.

@@

The data used in this analysis were collected from Spotify’s public API, which provides detailed metadata for songs. Each track is associated with multiple audio features and contextual information. The data was preprocessed to ensure consistency, with missing or erroneous values removed prior to analysis.

Valence (Outcome): Extracted directly from Spotify’s API, Valence is measured on a scale from 0 to 1, indicating the positivity or happiness conveyed by a song. This variable serves as the primary outcome of interest.

@@

Tempo: Captured in beats per minute (BPM), tempo describes the overall speed of a track. This feature was extracted directly from Spotify’s analysis tools.

@@

[Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.]

## Outcome variable
The distribution of Valence shows a relatively uniform spread, with some concentration near the mid-range. This reflects the diversity of emotional tones in the dataset. The histogram of Valence illustrates this distribution, where most tracks lie between moderate to high levels of positivity.
```{r}
#| label: fig-valence
#| fig-cap: Distribution of Valence
#| echo: false
#| warning: false
#| fig-align: center

@@

## Predictor variables
### Mean Temperature

The average temperature over time demonstrates a seasonal decline, ranging from approximately 25°C in the summer to below 10°C in late autumn. This variable adds a contextual layer, capturing external influences that may correlate with song Valence and listening behavior.
```{r}
#| label: fig-temp
#| fig-cap: Average temperature Over Time
#| echo: false
#| warning: false
#| fig-align: center
ggplot(analysis_data, aes(x = Date, y = `Mean Temp`)) +
  geom_line(color = "red") +
  labs(title = "Average Temperature Over Time", x = "Date", y = "Temperature (°C)") +
  theme_minimal()
```
### Tempo
The distribution of Tempo reveals a central peak around 120 BPM, with a wider spread for slower and faster songs. This suggests a balanced mix of tempos, with a slight preference for moderate-speed tracks. Tempo’s role in influencing Valence is analyzed alongside other musical features.

@@

```{r}
#| label: fig-tempo
#| fig-cap: Distribution of Tempo
#| echo: false
#| warning: false
#| fig-align: center
analysis_data %>%
  select(Danceability, Energy, Acousticness) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value, fill = Variable)) +
  geom_histogram(binwidth = 0.05, alpha = 0.6, position = "identity", color = "black") +
  facet_wrap(~Variable, scales = "free") +
  labs(
    title = "Distribution of Danceability, Energy, and Acousticness",
    x = "Value",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

@@

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

# Appendix {-}

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

\[Overview paragraph

@@

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....\]

@@

The dataset contains key variables spanning musical features, environmental data, and outcomes. Below is a summary of numerical variables:

```{r}
#| label: tbl-summary
#| tbl-cap: Summary Statistics
#| echo: false
#| warning: false
#| fig-align: center
summary_table <- analysis_data %>%
  reframe(
    Variable = c("Valence", "Mean Temp", "Danceability", "Energy", "Acousticness", "Tempo"),
    Mean = c(
      mean(Valence, na.rm = TRUE),
      mean(`Mean Temp`, na.rm = TRUE),
      mean(Danceability, na.rm = TRUE),
      mean(Energy, na.rm = TRUE),
      mean(Acousticness, na.rm = TRUE),
      mean(Tempo, na.rm = TRUE)
    ),
    SD = c(
      sd(Valence, na.rm = TRUE),
      sd(`Mean Temp`, na.rm = TRUE),
      sd(Danceability, na.rm = TRUE),
      sd(Energy, na.rm = TRUE),
      sd(Acousticness, na.rm = TRUE),
      sd(Tempo, na.rm = TRUE)
    ),
    Min = c(
      min(Valence, na.rm = TRUE),
      min(`Mean Temp`, na.rm = TRUE),
      min(Danceability, na.rm = TRUE),
      min(Energy, na.rm = TRUE),
      min(Acousticness, na.rm = TRUE),
      min(Tempo, na.rm = TRUE)
    ),
    Max = c(
      max(Valence, na.rm = TRUE),
      max(`Mean Temp`, na.rm = TRUE),
      max(Danceability, na.rm = TRUE),
      max(Energy, na.rm = TRUE),
      max(Acousticness, na.rm = TRUE),
      max(Tempo, na.rm = TRUE)
    )
  ) %>%
  mutate(
    Mean = round(Mean, 3),
    SD = round(SD, 3),
    Min = round(Min, 3),
    Max = round(Max, 3)
  )
kable(summary_table, format = "pipe")
```

These metrics reveal diverse song features and moderate variability in environmental temperature, offering rich explanatory potential for Valence.

@@

The data used in this analysis were collected from Spotify's public API, which provides detailed metadata for songs. Each track is associated with multiple audio features and contextual information. The data was preprocessed to ensure consistency, with missing or erroneous values removed prior to analysis.

Valence (Outcome): Extracted directly from Spotify's API, Valence is measured on a scale from 0 to 1, indicating the positivity or happiness conveyed by a song. This variable serves as the primary outcome of interest.

@@

Tempo: Captured in beats per minute (BPM), tempo describes the overall speed of a track. This feature was extracted directly from Spotify's analysis tools.

@@

\[Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.\]

## Outcome variable

The distribution of Valence shows a relatively uniform spread, with some concentration near the mid-range. This reflects the diversity of emotional tones in the dataset. The histogram of Valence illustrates this distribution, where most tracks lie between moderate to high levels of positivity.

```{r}
#| label: fig-valence
#| fig-cap: Distribution of Valence
#| echo: false
#| warning: false
#| fig-align: center

@@

## Predictor variables

### Mean Temperature

The average temperature over time demonstrates a seasonal decline, ranging from approximately 25°C in the summer to below 10°C in late autumn. This variable adds a contextual layer, capturing external influences that may correlate with song Valence and listening behavior.

```{r}
#| label: fig-temp
#| fig-cap: Average temperature Over Time
#| echo: false
#| warning: false
#| fig-align: center
ggplot(analysis_data, aes(x = Date, y = `Mean Temp`)) +
  geom_line(color = "red") +
  labs(title = "Average Temperature Over Time", x = "Date", y = "Temperature (°C)") +
  theme_minimal()
```

### Tempo

The distribution of Tempo reveals a central peak around 120 BPM, with a wider spread for slower and faster songs. This suggests a balanced mix of tempos, with a slight preference for moderate-speed tracks. Tempo's role in influencing Valence is analyzed alongside other musical features.

@@

```{r}
#| label: fig-diff
#| fig-cap: Distribution of Tempo
#| echo: false
#| warning: false
#| fig-align: center
analysis_data %>%
  select(Danceability, Energy, Acousticness) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value, fill = Variable)) +
  geom_histogram(binwidth = 0.05, alpha = 0.6, position = "identity", color = "black") +
  facet_wrap(~Variable, scales = "free") +
  labs(
    title = "Distribution of Danceability, Energy, and Acousticness",
    x = "Value",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

@@

## Model Structure

```{=text}
\begin{align}
y_i &= \beta_0 + \beta_1 (\text{Mean Temp}_i) + \beta_2 (\text{Artist}_i) + \beta_3 (\text{Danceability}_i) \\
&\quad + \beta_4 (\text{Energy}_i) + \beta_5 (\text{Acousticness}_i) + \beta_6 (\text{Tempo}_i) + \epsilon_i \\
\epsilon_i &\sim \text{Normal}(0, \sigma^2)
\end{align}
```
- $y_i$: 
- $\beta_0$:
- $\beta_1$: 
- ... [add beta's for the other predictors]
- ...
- $u_{\text{pollster}}$:
- $u_{\text{state}}$:

@@

```{r}
#| label: tbl-comp
#| fig-cap: Comparison of Model Performance
#| echo: false
#| warning: false
#| fig-align: center
comparison_data <- data.frame(
  Model = c("MLR 1", "MLR 2", "MLR 3"),
  `# Variables` = c(6, 5, 4),
  RMSE = c(0.0846, 0.0866, 0.2031),
  `R-Squared` = c(0.8676, 0.8600, -1.4266),
  check.names = FALSE
)

kable(comparison_data, format = "pipe")
```

@@

```

@@

```

@@

# Appendix {.unnumbered}

@@

```

@@

```
